Compositional equivalences based on Open pNets

### 2020-09-08 2016-Weak-FH-Bisim/WeakBisim.tex

In practice, we are dealing with finite representations of the (infinite) open automata. In \cite{hou:hal-02406098}, we defined a slightly modified definition of the ``coverage'' proof obligation, in the case of strong FH-Bisimulation. This modification is required to manage in a finite way all possible instantiations of an OT. In the case of weak FH-Bisimulation, the proof obligation from definition \ref{def-Weak-bisim} becomes:
      \begin{multline*}
\forall fv_{OT}.\, \Big\{ \Pred \land \Pred_{OT} \implies\\
 \bigvee_{x\in X}\!
  \left[\exists fv_{OT_x}.
    \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)\right]\Big\}
\end{multline*}

@@

\begin{definition}[Non-observability of silent actions for pNets]\label{def:non-obspNet}~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \set{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\True}\in \set{\symb{SV}}$ and 
\begin{equation*}
%\begin{split}
\forall \left(\SV{{(\alpha_i)}^{i\in I'}} 
{\alpha'} 
{e_b}\in \set{\symb{SV}}\right), %&
\forall i\in I'\cap J,\, \alpha_i=\tau \implies %\\
%& 
\alpha'=\tau ~\land~ %\\
%&
I'=\{i\}
%\end{split}
\end{equation*}
\end{definition}

@@

\begin{itemize}
  \item Their sets of active (non-silent) holes is the same: $J' = J_x = \{\texttt{P}\}$.
  \item Triple ($\text{b1},  100,  \texttt{b\_msg = s\_msg} \land \texttt{b\_ec = s\_ec}$) is in $\mathcal{R}$.
  \item The verification condition \\
    $\forall fv_{OT}.\, \{ \Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\!
   \left[\exists fv_{OT_x}.
  \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)\right]\}}$

@@

We are currently extending this work,  looking at  both further properties of FH-bisimulation, but also
the relations with existing equivalences on both closed and open systems. In particular, our model being significantly different from those considered in \cite{IngolfsdottirL:2001}, it
would be interesting to compare our "FH" family of bisimulations with the hierarchy of symbolic bisimulations from these authors.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.  
Naturally we are now working toward an implementation of weak FH-Bisimulation. 
The challenges here, in the context of our symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
Two main directions should be explored, either with an explicit construction of the weak transition, but this step in
itself may introduce non-termination, or a direct implementation of the weak bisimulation definition, without
constructing the weak automaton. We can expect that optimisations here in the symbolic setting would be quite different from the classical case.

@@

  \Pred_{OT_{c}}\subst{\Post_{OT_{b}}\shortotimes\Post_{OT_a}}}_{\Pred} \land  \underbrace{\alpha'\subst{\Post_1}}_{\alpha}=\underbrace{\alpha_{2 b} \subst{\Post_{OT_a}}}_{\alpha''} \land\\ \Pred_{s_{31},c}  \subst{\underbrace{\Post_3\shortotimes\Post_2\shortotimes\Post_1}_{\Post_{OT}}\uplus \underbrace{\Post_{OT_{c}}\shortotimes \Post_{OT_{b}}\shortotimes\Post_{OT_a}}_{\Post}})

>>>>

In practice, we are dealing with finite representations of the (infinite) open automata. In \cite{hou:hal-02406098}, we defined a slightly modified definition of the ``coverage'' proof obligation, in the case of strong FH-Bisimulation. This modification is required to manage in a finite way all possible instantiations of an OT. In the case of weak FH-Bisimulation, the proof obligation from Definition \ref{def-Weak-bisim} becomes:
      \begin{multline*}
\forall fv_{OT}.\, \Big\{ \Pred \land \Pred_{OT} \implies\\
 \bigvee_{x\in X}\!
  \left[\exists fv_{OT_x}.
    \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)\right]\Big\}
\end{multline*}

@@

\begin{definition}[Non-observability of silent actions for pNets]\label{def:non-obspNet}~\\
A pNet $\mylangle \pNet_i^{i\in I} , \Sort_j^{j\in J}, \set{\symb{SV}}\myrangle$
 \emph{cannot observe silent actions} if it verifies:\\ $\forall i\in I\uplus J.\, \SV{(i\mapsto \tau)}{\tau}{\True}\in \set{\symb{SV}}$ and 
\begin{equation*}
%\begin{split}
\forall \left(\SV{{(\alpha_i)}^{i\in I'}} 
{\alpha'} 
{e_b}\in \set{\symb{SV}}\right), %&
\forall i\in I'\cap J.\, \alpha_i=\tau \implies %\\
%& 
\alpha'=\tau ~\land~ %\\
%&
I'=\{i\}
%\end{split}
\end{equation*}
\end{definition}

@@

\begin{itemize}
  \item Their sets of active (non-silent) holes is the same: $J' = J_x = \{\texttt{P}\}$.
  \item Triple ($\text{b1},  100,  \texttt{b\_msg = s\_msg} \land \texttt{b\_ec = s\_ec}$) is in $\mathcal{R}$.
  \item The verification condition \\
    $\forall fv_{OT}.\, \{ \Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \displaystyle{\bigvee_{x\in X}\!
   \left[\exists fv_{OT_x}.
  \left( \forall j\in J_x. \vis{\beta_j}\!=\!\gamma_{jx}\! \land\! \Pred_{OT_x}
     \!\land\! \alpha\!=\!\alpha_x\! \land\!  
     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\!\Post_{OT_x}}\right)\right]\}}$

@@

We are currently extending this work,  looking at further properties of FH-bisimulation, but also
the relations with existing equivalences on both closed and open systems. In particular, our model being significantly different from those considered in \cite{IngolfsdottirL:2001}, it
would be interesting to compare our "FH" family of bisimulations with the hierarchy of symbolic bisimulations from these authors.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have developed tool support for computing the
symbolic semantics in term of open automata \cite{QBMZ-AVOCS18}, and have
developed algorithms to check strong FH-bisimulation \cite{hou:hal-02406098}.  
Naturally we are now working toward an implementation of weak FH-Bisimulation. 
The challenges here, in the context of our symbolic systems, is not so much algorithmic complexity,
as was the case with classical weak bisimulation on finite models, but decidability and termination.
Two main directions should be explored, either with an explicit construction of the weak transition, but this step in
itself may introduce non-termination, or a direct implementation of the weak bisimulation definition, without
constructing the weak automaton. We can expect that optimisations here in the symbolic setting would be quite different from the classical case.

@@

  \Pred_{OT_{c}}\subst{\Post_{OT_{b}}\shortotimes\Post_{OT_a}}}_{\Pred} \land  \underbrace{\alpha'\subst{\Post_1}}_{\alpha}=\underbrace{\alpha_{2 b} \subst{\Post_{OT_a}}}_{\alpha''} \\\land \Pred_{s_{31},c}  \subst{\underbrace{\Post_3\shortotimes\Post_2\shortotimes\Post_1}_{\Post_{OT}}\uplus \underbrace{\Post_{OT_{c}}\shortotimes \Post_{OT_{b}}\shortotimes\Post_{OT_a}}_{\Post}})
